---
title: "Getting Started with replr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with replr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

`replr` provides isolated REPL (Read-Eval-Print Loop) functionality for R, allowing you to execute R code in separate worker processes. This vignette demonstrates the core features and usage patterns.

## R6 Class Interface

The primary interface for `replr` is the `RREPLSession` R6 class:

```{r basic-usage, eval = FALSE}
library(replr)

# Create a new isolated R session
session <- RREPLSession$new()

# Execute R code in the isolated process
result <- session$execute("2 + 2")
print(result$result$output)  # [1] "[1] 4"

# Check session status
session$is_alive()  # TRUE
session$port        # Port number
session$pid         # Process ID

# Stop the session (or it will auto-cleanup on garbage collection)
session$stop()
```

## Error Handling

Workers are resilient and continue processing after errors:

```{r error-handling, eval = FALSE}
library(replr)

session <- RREPLSession$new()

# Handle syntax errors
result1 <- session$execute("1 +")
print(result1$status)         # [1] "error"
print(result1$result$errors)  # [1] "Worker error: <text>:2:0: unexpected end of input\n1: 1 +\n   ^"

# Handle runtime errors
result2 <- session$execute("stop('Custom error')")
print(result2$status)         # [1] "error"
print(result2$result$errors)  # [1] "Worker error: Custom error"

# Worker continues after errors
result3 <- session$execute("3 * 4")
print(result3$status)         # [1] "success" - worker survived!

# Code that generates warnings
result4 <- session$execute("
  warning('This is a warning')
  42
")

print(result4$status)             # [1] "success"
print(result4$result$warnings)    # [1] "This is a warning"
print(result4$result$output)      # [1] "[1] 42"

session$stop()
```

## Debug Logging

Enable debug logging to see detailed information about worker processes:

```{r debug-logging, eval = FALSE}
library(replr)

# Enable debug logging
enable_debug(TRUE)

# Now see output of what is happening
session <- RREPLSession$new()  # Shows startup debug messages

result <- session$execute("sqrt(16)")  # Shows communication debug

# Get logs programmatically
logs <- session$get_debug_logs()
cat("Worker generated", length(logs), "debug messages\n")

session$stop()  # Shows cleanup debug

# Disable debug logging
enable_debug(FALSE)

# Check current debug status
debug_status()
```

## Multiple Concurrent Sessions

You can run multiple isolated sessions simultaneously:

```{r multiple-sessions, eval = FALSE}
library(replr)

# Start multiple isolated sessions
session1 <- RREPLSession$new()
session2 <- RREPLSession$new()

# They run on different ports and are completely isolated
print(session1$port)  # e.g., 5555
print(session2$port)  # e.g., 5556

# Set different variables in each session
session1$execute("my_var <- 'Session 1'")
session2$execute("my_var <- 'Session 2'")

# Verify isolation - each session only sees its own variables
result1 <- session1$execute("my_var")
result2 <- session2$execute("my_var")

print(result1$result$output)  # [1] "[1] \"Session 1\""
print(result2$result$output)  # [1] "[1] \"Session 2\""

# Clean up both sessions
session1$stop()
session2$stop()
```

## Advanced Usage with Timeouts

Control execution timeouts for long-running operations:

```{r timeouts, eval = FALSE}
library(replr)

session <- RREPLSession$new()

# Set custom timeout for long-running operations
result <- session$execute(
  "Sys.sleep(2); 'Completed'",
  timeout = 5  # 5 second timeout
)
print(result$result$output)  # [1] "[1] \"Completed\""

# This would timeout:
# result <- session$execute("Sys.sleep(10)", timeout = 2)

session$stop()
```

## Next Steps

- Learn about [Docker container support](https://github.com/pkrusche/replr#docker-container-support) for enhanced security
- Explore [ellmer integration](https://github.com/pkrusche/replr#ellmer-tools-for-llm-agents) for LLM agent capabilities
- Check out example scripts in `inst/examples/`
